Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-20 manuscript/manuscript.qmd
Remove unncessary files

---
---

@@

library(knitr)
library(kableExtra)
library(ggplot2)
library(gt)
library(gtExtras)
library(patchwork)
library(english)
library(gtExtras)
library(marginaleffects)

@@

Building a mental lexicon is a major achievement in the development of an infant: by storing representations of how familiar words sound and what they mean, an infant is able to make sense of their linguistic input. The foundations of an initial lexicon are in place before the end the first year of life [@tincoff1999some; @halle1994emergence; @vihman2004cross; @parise2012electrophysiological; @bergelson20126; @bergelson2015early]. This initial lexicon consists of only a few items; mainly words for people, interjections, body parts, and food [@tincoff2012six; @tardif2008baby], but it undergoes rapid growth during the second year of life [@ganger2004reexamining; @goldfield1990early; @bloom2002children; @mcmurray2007defusing; @bergelson2020comprehension]. According to parental reports, the average 15-month-old infant already understands more than 100 words, and by two years of age, they understand more than 400 [@frank2021variability]. This accelerated lexical developmental is reflected in infants' trajectories of word recognition: infants recognise familiar words faster and more efficiently as they approach their second birthday [@fernald1998rapid; @fernald2001half; @hurtado2007spoken]. Despite being exposed to a more complex linguistic input, bilinguals show equivalent trajectories of word acquisition and word recognition to their monolingual peers' [@vihman2007onset; @legacy2018vocabulary; @hoff2012dual; @de2014bilingual; @pearson1994patterns; @byers-heinlein2023sometimes; @bialystok2009bilingualism]. This is a remarkable deed for two reasons. First, bilingual infants receive a relative impoverished quantitative linguistic input in each of their languages, compared to monolinguals [@costa2014does; @cattani2014much; @thordardottir2011relationship]. Second, they face a more complex referential context, as they often learn two labels for each referent, one in each language [@au1990principle; @tsui2022translation; @bilson2015semantic; @de2006early]. The mechanisms that allow bilingual' trajectories of lexical developmental to keep up with monolinguals' are still unclear.

Previous studies have pointed to cross-language dynamics as a candidate mechanism boosting lexical acquisition in bilinguals [@floccia2018vocabulary; @blom2020cross; @gampe2021does]. In particular, lexical similarity between the two languages has been associated with larger vocabulary sizes in bilingual toddlers. @floccia2018vocabulary collected parental reports of vocabulary data from a sample of 367 bilingual children living in the United Kingdom, who were learning English and an additional language (out of a diverse pool of 13 languages). The authors then used an edit distance-based metric of phonological similarity to compute the average lexical similarity between the translation pairs in English and each of the additional languages. English and Dutch shared the highest lexical similarity, while English and Mandarin shared the lowest. Children's vocabulary sizes in the additional language was positively associated with the amount of language similarity between their two languages. For instance, children learning English and Dutch showed larger vocabulary sizes than those learning English and Mandarin. The authors suggested that the acquisition of words in the additional language might be facilitated by their cognate status (i.e., being phonologically similar to their translation equivalent). If this is the case, larger vocabulary sizes might then be expected in bilinguals learning two languages sharing a high proportion of cognates. This would be consistent with available evidence of an earlier acquisition of cognate words [@bosch2014first; @mitchell2022cognates; @garcia2023cognate].

The facilitative effect of cognateness is in line with the language non-selective account of bilingual lexical access. This account proposes that bilinguals activate both languages in parallel, even during monolingual situations, as suggested by robust experimental evidence [@marian1999activation; @groot1992determinants; @schwartz2007reading; @dijkstra1999recognition; @dufour1995matching; @dijkstra2010cross]. For instance @costa2000cognate presented Catalan-Spanish bilinguals with a series of pictures of familiar objects. Participants were asked to name each picture in Catalan, their dominant language. Unbeknownst to participants, the authors manipulated the cognate status pictures' labels in Catalan and Spanish. In half of the trials, the labels associated to the pictures were cognates (i.e., phonologically similar translation equivalents, e.g., *gat*-*gat* [*cat*]) in Catalan and Spanish, whereas in the other half of the trials the labels were non-cognates (e.g., *taula*-*mesa* [*table*]). Participants named pictures faster in cognate trials than in non-cognate trials. Critically, Spanish monolinguals---who completed the same task in Spanish---showed equivalent naming times in both conditions. These results revealed that bilinguals activated their Spanish phonology, despite performing the naming task exclusively in Catalan: the visual recognition of the presented pictures led to the parallel activation of its associated phonological forms in both languages, which influenced the subsequent dynamics of word production.

Parallel activation has also been reported in the developing lexicon [@poarch2012cross; @bosma2020cognate; @singh2014one; @jardak2019labels; @von2019impact; @floccia2020translation; @bosma2019longitudinal]. For instance, @von2012language found suggestive evidence of cross-language phonological priming in a sample of 20 German-English bilinguals aged 21 to 43 months. In their experimental task, each trial begun with the auditory presentation of an English prime word, followed by a target German label and a pair of target and distractor pictures. The authors registered participants' target picture looking (interpreted as a proxy of target word recognition) under three conditions. In the *phonological priming* condition, the auditorily presented prime and target labels were phonologically related (e.g., *slide*-*Klide* [*dress*]). In the *priming through translation* condition, prime and target were phonologically related through translation: the auditorily presented prime label did *not* overlap with the target label (e.g., *leg*-*Stein* [*stone*]), but its German translation did [*Bein*]. In the *unrelated* condition, prime and target labels were phonologically unrelated in both languages. Interestingly, participants showed a weaker target preference in priming through translation trials, compared to unrelated trials. This was interpreted evidence of the lexical representations of the prime labels being activated in both English and German, leading to interference effects in later target recognition. Previous studies had reported similar cross-language priming effects, but the conclusion that such effects occurred at the lexical level, as opposed to being the result of mere acoustic similarity, was nonetheless unwarranted. In studies where the auditorily presented prime label was phonologically overlapped with the target label [e.g., @marian1999activation], priming effects could be explained by sub-lexical facilitation across phonological segments. This might explain the facilitation effect found in *phonological priming* trials in @von2012language, in which English prime and German target labels shared overt phonological overlap. In contrast, the interference effect found in *priming through translation* trials points to cross-language competition at the lexical level. Since participants' target recognition was modulated by the phonology of the German translation of the prime (which participants did *not* hear), cross-language priming can be explained by the lexical co-activation of the English *and* German prime labels.

@von2012language extended the available knowledge on the developing bilingual lexicon, but some methodological challenges prevent drawing strong conclusions from their findings. First, bilingual participants' quantitative exposure to English as a second language was lower than what is conventionally considered as bilingualism [20-50% of cumulative exposure, @byers-heinlein2021multilab; @rocha-hidalgo2023defining]^[Except for four participants who were also exposed to English at home, participants in @von2012language received their English input exclusively at school from a German native speaker. Participants' exposure to each language ranged between three and four hours per day. This implies that participants' exposure to English consisted of an average of 3-4 hours per day, during the time they attended daycare (reported as an average 12.35 months). Assuming that infants were exposed to a daily average of 12 total hours of speech, and that they attended daycare five days per week, the estimated cumulative to English in the life of an average child in the sample was between `r scales::percent((3*5*4*12.35)/(12*7*4*21), accuracy = 0.01)` and `r scales::percent((4*5*4*12.35)/(12*7*4*21), accuracy = 0.01)` for the 21 month-olds (the youngest in the sample), and between `r scales::percent((3*5*4*12.35)/(12*7*4*43), accuracy = 0.01)` and `r scales::percent((4*5*4*12.35)/(12*7*4*43), accuracy = 0.01)` for the 43-month-olds. These percentages were calculated as:
$$\frac{\text{Total English input}}{ \text{Total English + German input}} = \frac{3 \text{ or } 4 \text{ hours/day} \times 5 \text{ days/week } \times 4\text{ weeks/month }  \times 21 \text{ or } 43 \text{ months of age}}{12 \text{ hours/day} \times 7 \text{ days/week } \times 4\text{ weeks/month }  \times 21 \text{ or } 43 \text{ months of age}}$$.]. Results should therefore be taken with caution when generalised to other populations of bilinguals. A second consideration is the presence of cognates in a substantial proportion of the trials in the *phonological priming* and *priming thorugh translation* conditions. Because of this, some prime labels in *phonological priming* trials shared substantial cross-language phonological overlap thought translation with the target words. Conversely, some trials in the *phonological priming through translation* shared within-language overlap with the target. Third, the observed pattern of results was dependent on the inclusion of a single item, *Buch* [*book*, in German]. When this item was included, monolinguals showed a similar interference effect to bilinguals' in the *priming through translation* condition, in which only bilinguals with knowledge of the translation should show an effect. Fourth, data from monolinguals and bilinguals were not compared directly (analyses were conducted separately for both groups). Differences in statistical significance between monolinguals and bilinguals across conditions do not necessarily guarantee significant differences between the two groups. As a consequence, whether the interference effects observed in the *priming through translation* condition is exclusive to bilinguals (which would suggest that the priming effect were not due to the intrinsic properties of the stimuli) or not is unclear. In summary, evidence of cross-language activation at the lexical level in infants is, to date, inconclusive.

In the present study, we use an extension of the implicit naming task [@mani2010infant; @mani2011phonological] to put the language non-selectivity in the developing lexicon to a more severe test. Previous studies have successfully used this task to investigate priming effects in word recognition in monolingual infants and bilingual adults. In the original implementation of the task, @mani2010infant primed infants with a picture presented in silence. At test, a target-distractor picture pair were presented side-by-side. Finally, the target picture's label was presented auditorily. The authors manipulated the phonological overlap between the prime and the target labels, so that in half of the trials both labels were phonologically related, sharing phonological onset (*cat*-*cup*), or phonologically unrelated (*ball*-*comb*). Prime and target-distractor pairs were semantically unrelated. Eighteen-month-old infants showed a stronger looking preference for the target picture after phonologically related prime pictures, compared to after phonologically unrelated primes. This suggests that infants implicitly named prime pictures despite such pictures having been presented in silence, and that the phonology of the resulting label interacted with the subsequent auditory recognition of the target label.

@von2014bilinguals adapted this paradigm to investigate cross-language activation in bilingual adults. The authors presented German-English bilinguals with 120 prime-target pairs. Primes were presented as familiar pictures in silence. After prime picture offset, target words were presented auditorily. Participants' eletrophysiological response was recorded, and N400 evoked potentials to the auditory target label was measured. The authors manipulated the phonological relationship between the prime and target word-forms within participants' L1 and across L1 and L2. When prime and target were phonologically identical (*Affe*-*Affe*, German for *monkey*), or similar (*Fahne*-*Sahne*, German for *flag* and *ice-cream*, respectively) within L1, participants' showed a reduced N400 amplitude, compared to when prime and target were phonologically unrelated (*Messer*-*Seil*, German for *knife* and *rope*). Critically, a similar effect was found when prime and target were phonologically related through translation (*Rustsche*-*Kleid*, German for *slide* and *dress*). This suggests that participants activated prime labels in both languages in parallel, and that both labels impacted the dynamics of target words recognition. In summary, the implicit naming paradigm has been successful used in the past to investigate the emergence of phonological priming effects in monolingual infants, and to investigate cross-language priming effects in bilingual adults.

The implicit naming paradigm provides several advantages over the priming through translation task. In the priming through translation task, primes and targets are auditorily presented in both languages in the same trial. This presents participants with a bilingual context that might lead to a higher baseline activation of lexical representations of both languages, contrary to a context in which parallel activation results from genuine cross-language activation [@grosjean1997bilingual]. Also, when prime words are auditorily presented, the task imposes an asymmetry between the activation dynamics of the lexical representation of the prime in each language: while participants hear prime labels in one language, the activation of prime labels in the other must occur via cascaded activation across semantic links with their translations. This might lead to a lower strength, and higher latency of the activation of prime labels in one language, compared to the presented labels in the other. In contrast, the implicit naming paradigm provides participants with an exclusively monolingual context in which they perform the task in only one of their languages. At the same time, the silent presentation of prime pictures allows participants to activate prime labels in both languages in a truly parallel fashion. If differences in task performance are found associated with covert manipulation of cross-linguistic properties of prime labels, such differences may be more conclusively linked to the parallel activation of lexical representations in both languages.

In the present investigation, we manipulated both the phonological overlap between the prime word and target word and the cognate status of the prime word, to test the whether bilingual toddlers implicitly name familiar pictures in both languages. If this is the case, phonological priming effects should be larger when the prime label in both languages shares phonological overlap with the target words (*Related/Cognate* condition), compared to when only the label of the prime in one language overlaps phonologically with the target (*Related/Non-cognate*). We also also included a control *Unrelated* condition in which regardless of the cognate status of the prime word, neither of the prime labels in any language overlapped phonologically with the target word. We registered participants' target looking during target-distractor presentation, and that, if within-language priming effects take place, differences in target looking between *Unrelated* and *Related/Non-cognate* trials should be observed, as previously reported [@von2012language, @mani2010infant; @mani2011phonological]. We further predicted cross-language priming effects, revealed as differences in target looking between *Related/Non-cognate* and *Related/Cognate* trials. This effect should be stronger in bilinguals than monolinguals, if not completely absent in the latter, who should be indifferent to the phonological properties of the stimuli in the other language.

Following previous studies associating the emergence of phonological links between lexical items to the size of children's vocabulary [@mani2011phonological; @chow2017spoken], we also compared the explanatory power of two indices of receptive vocabulary size against participants' age: dominant-language vocabulary size and total vocabulary size. If within-language or cross-language phonological associations in the lexicon emerge as a function of vocabulary growth, the size of the differences between conditions should larger in participants with larger vocabulary sizes. A model including participants' dominant-language or total vocabulary size as a predictor should then show better performance than a model that instead includes participants age. In the case of bilinguals, if the emergence of cross-language priming effects is associated with the growth of their vocabulary is *both* languages, a model including total vocabulary size as a predictor should perform better than a model that includes dominant-language vocabulary size. The opposite pattern would indicate that cross-language priming effects emerge as a function of the growth of dominant-language vocabulary.

Testing bilingual populations at early ages presents the limitation that the vocabulary size of participants' non-dominant language is relatively small, which substantially constrains the number of items that can be used to as stimuli in the task, as well as the number of trials that can be constructed to generate stimuli lists. In order to circumvent the problem of limited vocabulary knowledge in the non-dominant language, tested participants only in their dominant language, therefore directing our research to the question of how possessing a second language affects processing of the dominant one in bilingual toddlers [@costa2014does provide a fuller description of the advantages of such approach].

@@

@@

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/ckydb/)) and GitHub ([https://github.com/gongcastro/cognate-priming](https://github.com/gongcastro/cognate-priming)) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I. Before every testing session, caregivers were asked to read and sign an informed consent form, and were given a token of appreciation at the end of it.

@@

We tested the differences between the conditions of interest in two ways. First, we examined the posterior distribution of the regression coefficients of the linear predictors in model $\mathcal{M}_0$ (see @eq-model). We assessed the practical relevance of the coefficients following @kruschke2018bayesian. We specified a region of practical equivalence (ROPE) from `r rope_interval[1]` to +`r rope_interval[2]`, in the logit scale. This region indicates the range of values that we considered equivalent to zero. We then summarised the posterior distribution of each regression coefficient with the 95% highest density interval (HDI). This interval contains the true value of this coefficient with 95% probability, given the data. Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, noted as $p(\text{ROPE})$, which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero). For example, $p(\text{ROPE})=.80$ indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero. Second, we examined the differences between the priming conditions across the time course of the trial. We generated marginal posterior predictions from the model for the three conditions, for monolinguals and monolinguals. These predictions included the non-linear relationship between target looking and the time course of the trial. We then computed the difference between the marginal posterior predictions for the *Unrelated* and *Related/Non-cognate* conditions, and between those for the *Related/Non-cognate* and *Related/Cognate* conditions, at each time bin. Finally, we compared the 95% HDI of the differences against the [-0.1, +0.1] ROPE.

@@

In this study, we conducted a conceptual replication of @mani2010infant and @mani2011phonological. Using the same task from Study 1, we tested a group of monolinguals aged 20 to 32 months being exclusively raised in British English speaking homes, the same population tested in the original studies. For consistency with Study 1, we also manipulated the phonological relationship between the prime and target labels at word onset, as well as the cognate status of the prime label. Evidence from implicit naming of prime images would be revealed by differences in target looking or its time course across the trials between *Unrelated* trials and *Related/Non-cognate*, and between *Related/Non-cognate* and *Related/Cognate* trials. Given that participants in this study are monolinguals---and contrary to the monolinguals in Barcelona, have received no exposure to any other language---we should not find the cognate status of the prime labels to exert any effect on subsequent target recognition.

>>>>

@@

library(knitr)
library(kableExtra)
library(ggplot2)
library(gt)
library(gtExtras)
library(patchwork)
library(english)
library(gtExtras)

@@

Building a mental lexicon is a major achievement in the development of an infant: by storing representations of how familiar words sound and what they mean, an infant is able to make sense of their linguistic input. The foundations of an initial lexicon are in place before the end the first year of life [@tincoff1999some; @halle1994emergence; @vihman2004cross; @parise2012electrophysiological; @bergelson20126; @bergelson2015early]. This initial lexicon consists of only a few items; mainly words for people, interjections, body parts, and food [@tincoff2012six; @tardif2008baby], but it undergoes rapid growth during the second year of life [@ganger2004reexamining; @goldfield1990early; @bloom2002children; @mcmurray2007defusing; @bergelson2020comprehension]. According to parental reports, the average 15-month-old infant already understands more than 100 words, and by two years of age, they understand more than 400 [@frank2021variability]. This accelerated lexical developmental is reflected in infants' trajectories of word recognition: infants recognise familiar words faster and more efficiently as they approach their second birthday [@fernald1998rapid; @fernald2001half; @hurtado2007spoken]. Despite being exposed to a more complex linguistic input, bilinguals show equivalent trajectories of word acquisition and word recognition to their monolingual peers' [@vihman2007onset; @legacy2018vocabulary; @hoff2012dual; @de2014bilingual; @pearson1994patterns; @byers-heinlein2023sometimes; @bialystok2009bilingualism]. This is a remarkable deed for two reasons. First, bilingual infants receive a relatively reduced linguistic input in each of their languages, compared to monolinguals [@costa2014does; @cattani2014much; @thordardottir2011relationship]. Second, they face a more complex referential context: they often learn two labels for each referent (one in each language), which additionally may not be direct translations of each other [@au1990principle; @tsui2022translation; @bilson2015semantic; @de2006early]. The mechanisms that allow bilingual' trajectories of lexical developmental to keep up with monolinguals' are still unclear.

Previous studies have pointed to the similarity between the two languages of exposure as a facilitator of lexical acquisition in bilinguals [@floccia2018vocabulary; @blom2020cross; @gampe2021does]. @floccia2018vocabulary reported larger vocabulary sizes in bilingual toddlers leaning two languages that shared high lexical similarity. The authors collected parental reports of vocabulary data from a sample of 367 bilingual children living in the United Kingdom, who were learning English and an additional language (out of a diverse pool of 13 languages). The authors then calculated the average phonolexical similarity between English and each of the additional languages. English and Dutch shared the highest similarity, while English and Mandarin shared the lowest. Overall, children's vocabulary sizes in the additional language was positively associated with the amount of language similarity between their two languages. For instance, English-Dutch bilinguals showed larger vocabulary sizes in Dutch than English-Mandarin bilinguals did in Mandarin. The authors suggested that the acquisition of words in the additional language might be facilitated by their cognate status (i.e., being phonologically similar to their translation equivalent). If this is the case, larger vocabulary sizes might then be expected in bilinguals learning two languages sharing a high proportion of cognates. This would be consistent with available evidence of an earlier acquisition of cognate words [@bosch2014first; @mitchell2022cognates; @garcia2023cognate; @schelletter2002effect].

The facilitation effect of cognateness is in line with the language non-selective account of bilingual lexical access. This account proposes that bilinguals activate both languages in parallel, even during monolingual situations. In adults, there is robust evidence in favour of this language non-selective account of lexical access [@marian1999activation; @groot1992determinants; @schwartz2007reading; @dijkstra1999recognition; @dufour1995matching; @dijkstra2010cross; @spivey1999cross]. @costa2000cognate presented highly-proficient Catalan-Spanish bilinguals with a series of pictures of familiar objects. Participants were asked to name each picture in Spanish. Unbeknownst to participants, the authors manipulated the cognate status pictures' labels in Catalan and Spanish. In half of the trials, the labels associated to the pictures were cognates (e.g., *cat*-*gat* [cat]), whereas in the other half of the trials the labels were non-cognates (e.g., *taula*-*mesa* [table]). Participants named pictures faster in cognate trials than in non-cognate trials. Spanish monolinguals showed equivalent naming times in both conditions. These results revealed that bilinguals activated their Catalan phonology, despite performing the naming task exclusively in Spanish: the visual recognition of the presented pictures led to the parallel activation of its associated phonological forms in both languages, which influenced the subsequent dynamics of word production.

Parallel activation has also been reported in the developing lexicon [@poarch2012cross; @bosma2020cognate; @singh2014one; @jardak2019labels; @von2019impact; @floccia2020translation; @bosma2019longitudinal]. @von2012language found evidence of cross-language phonological priming in a sample of 20 German-English bilinguals aged 21 to 43 months. In their experimental task, each trial begun with the auditory presentation of an English prime word, followed by a target word in German, and a pair of target and distractor pictures. The authors recorded participants' target picture looking as a measure of target word recognition. The authors manipulated the cross-linguistic phonological overlap between the prime and the target labels. In a *priming through translation* condition, the English prime labels (leg) did not overlap with the German target labels (*Stein* [stone]), but with their German translations (*Bein*) did. In the *unrelated* condition, prime and target labels were not phonologically related in either German or English. If participants accessed their lexicon in a language non-selective way, the auditory presentation of the prime label in English should lead to the co-activation of its German translation. If this is the case, target word recognition should be interfered by the prior activation of a phonologically related German prime label. Under this hypothesis, the authors anticipated an delayed target looking in priming through translation trials, compared to unrelated trials. The results supported this hypothesis. In spite of the relevance of @von2012language study, some methodological issues deserve some consideration. First, for most participants, exposure to English (the less prevalent language) was lower than the minimal amount conventionally considered the threshold for bilingual exposure [@byers-heinlein2021multilab; @rocha-hidalgo2023defining]. Second, some of the prime labels in the priming through translation condition were cognates. If both English and German labels overlap phonologically with the German target label, priming effects can be explained by interference between words from the same language, as opposed to cross-language interference. Third---and most critically---participants were exposed to both English and German word in a by-trial basis. This creates a context in which interference effects may not have arised from the competition between the prime translation and the target words, but between the target word and any other word in the other language. Paradigms in which the experimental task is conducted exclusively in one language, while cross-linguistic features are covertly manipulated, offer a methodologically stronger basis for studying language non-selectivity in the developing lexicon [@grosjean1997bilingual].

@mani2010infant designed an implicit naming task, in which primes consisted of pictures presented in silence, instead of auditory labels. In each trial, English monolingual infants were first presented with pictures of familiar objects for 1,500 ms. Then, a target-distractor picture pair was presented for 2,000 ms, and then the auditory label of the target picture was presented. Post-naming target looking was recorded for another 2,000 ms until the end of the trial, as a measure of target word recognition. The authors manipulated the phonological overlap between the prime and the target labels, so that in half of the trials both labels were phonologically related, sharing phonological onset (cat-cup), or phonologically unrelated (ball-comb). Prime, target and distractor were unrelated otherwise. At 18 months of age, participants showed a stronger looking preference for the target pictures after phonologically related prime pictures, compared to after phonologically unrelated primes. Since the prime pictures were presented in silence, their results suggested that infants implicitly named the prime pictures, and that the phonology of the resulting word interacted with the subsequent recognition of the auditory target word. Later, @mani2011phonological tested 21-month-old infants in the same task. This time, priming effects were observed in the opposite direction: when prime and target labels were phonologically related, participants showed significantly weaker target looking preference, compared to unrelated trials. The size of this interference effect was associated to participants' vocabulary size, and to the cohort size of the prime label. The authors interpreted this finding as indicating a developmental shift. At 18 months, participants' word recognition might have experienced a sub-lexical facilitation effect, in which the prior activation of the shared phonological onset between the prime and target labels boosted word recognition. In older participants, the lexicon might have reached a critical size at which the recognition of the target was delayed by the activation of its phonological cohort.

The implicit naming paradigm provides an ideal experimental paradigm study the developing bilingual lexicon. By covertly manipulating the cross-linguistic relationship between the prime and target labels, parallel activation can be tested while participants are presented with auditory stimuli (target labels) exclusively in one of their languages [see @von2014bilinguals for a similar approach in bilingual adults]. Capitalizing on the language non-selective account of lexical access, we exploited the implicit naming to investigate the mechanisms behind the emergence of phonological priming effects in the bilingual developing lexicon. We tested a cohort of monolingual and bilingual infants learning Catalan and Spanish between 20 and 32 months of age. This allowed us to compare the performance of participants with differing vocabulary sizes in the word recognition task. In order to circumvent the problem of limited vocabulary knowledge in the non-dominant language, we tested participants only in their dominant language [@costa2014does].

Following @mani2010infants, each trial in the task started with the silent presentation of a prime picture. Both monolingual and bilingual infants were expected to implicitly name the prime picture. According to the language non-selective hypothesis of lexical access, bilinguals should generate two labels for the prime picture, one in each language. To test this prediction, we manipulated the phonological similarity between the prime and the target words in both languages (see @fig-task). In *Related/Non-cognate* trials, prime and target labels shared phonological onset in only the language of test. For instance an infant tested in Catalan would be presented with a chair as prime picture ($\textbf{c}\text{adira}_{\text{CAT}}/\text{silla}_{\text{SPA}}$) and with $\textbf{c}\text{ullera}_{\text{CAT}}$ [spoon] as target label. In line with @mani2011phonological, we anticipated that the phonological overlap between prime and target should modulate target word recognition in both monolinguals and bilinguals. This should be reflected in a delayed target looking preference, compared to *Unrelated* trials, in which prime and target did not share phonological onset. In *Related/Cognate* trials, the prime shared phonological onset with the target in both languages. For instance, the same infants tested in Catalan would be presented with a car as prime picture ($\textbf{c}\text{otxe}_{\text{CAT}}/\textbf{c}\text{oche}_{\text{SPA}}$) and with $\textbf{c}\text{ullera}_{\text{CAT}}$ [spoon] as target label. In bilinguals, parallel activation of the prime in both languages should increase the cohort of the target word, leading to stronger interference effects in this condition, compared to *Related/Non-cognate* and *Unrelated* trials.

![Predicted priming effects (or their absence) in the *Related/Cognate*, *Related/Non-cognate*, and *Unrelated* conditions, with examples for a participant tested in Catalan. Words represent lexical representations. Lexical representations of the task-relevant language (Catalan) are depicted inside grey boxes. Solid arrows indicate within-language priming effects, and dashed lines indicate cross-language priming effects. In *Related/Cognate* (A) and *Related/Non-cognate* (B) trials, the recognition of the Catalan target word \textipa{kuˈʎe.R@} [spoon] is predicted to be modulated by the prior activation of the prime label in Catalan. In *Related/Cognate* trials, the parallel activation of the prime label in Spanish is predicted to increase the strength of the priming effect.](_assets/img/hypotheses.png){fig-align="center"}

In line with previous studies in monolinguals, we further predicted that the strength of the lexical interference effects in the *Related/Non-cognate* and *Related/Cognate* conditions would be associated to participants' vocabulary size. Target word recognition should be delayed by the activation of a larger cohort of phonologically related words [@mani2011phonological; @chow2018phonological]. In the case of bilinguals, two measures of vocabulary size are relevant: *total vocabulary size* and *dominant-language vocabulary size*. Total vocabulary size refers to the number of words that a child has acquired, regardless of the language they belong to. Dominant-language vocabulary size (L1 vocabulary size, henceforth) refers to the number of acquired words in the dominant language of the child (in which they were tested), the language in which they know more words. If the size of interference effects are more tightly associated to bilinguals’ total vocabulary size, it would support the assumption that lexical representations from both languages enter a unique cohort in bilinguals.

Because of the short-lived effects of cross-language activation on lexical processing, and to maximise the probability of detecting priming effects, we introduced a change in the sequence of the trials relative to the original implementation by @mani2010infant. We presented target auditory labels inmediately after the offset of the prime picture, and before the onset of the target and distractor pictures. By presenting prime pictures and target auditory labels closer in time, implicit naming of the prime picture should be more likely to influence the recognition of the target word. To test the effects of this methodological change, we first run a control experiment, Study 1, in which we tested a group of same-aged English monolinguals.

@@

In this study, we conducted a conceptual replication of @mani2010infant and @mani2011phonological. We tested a group of English monolinguals aged 20 to 32 months, the same population tested in the original studies. For consistency with Study 1, we also manipulated the phonological relationship between the prime and target labels at word onset, as well as the cognate status of the prime label. Evidence from implicit naming of prime images would be revealed by differences in target looking or its time course across the trials between *Unrelated* trials and *Related/Non-cognate*, and between *Related/Non-cognate* and *Related/Cognate* trials. Given that participants in this study are monolinguals---and contrary to the monolinguals in Barcelona, have received no exposure to any other language---we should not find the cognate status of the prime labels to exert any effect on subsequent target recognition.

@@

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/ckydb/)) and GitHub (<https://github.com/gongcastro/cognate-priming>) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I. Before every testing session, caregivers were asked to read and sign an informed consent form, and were given a token of appreciation at the end of it.

We collected data from `r n_participants_total` children (`r n_participants_sex$n[1]` female, `r n_participants_sex$n[2]` male, with `r english::words(n_participants_sex$n[3])` additional participants' sex not being reported; Age: *Mean* = `r round(ages$age_mean, 2)` months, *SD* = `r round(ages$age_sd, 2)`, *Range* = \[`r round(ages$age_min, 2)`-`r round(ages$age_max, 2)`\]), living in the Oxfordshire area (United Kingdom). Participants were tested at the Oxford BabyLab at the University of Oxford. Families were recruited from maternity rooms in private hospitals and social media, and contacted via phone when the child's age spanned between 20 and 32 months. From the `r n_participants_total` children that participated, `r n_participants_sessions$session_1` participated once, and `r n_participants_sessions$session_2` participated twice. Recurrent participants were tested with at least `r round(time_diffs, 2)$diff_min` months of difference. We gathered a total of `r n_sessions_total$n` testing sessions. All participants were being raised in exclusively British English monolingual homes.

```{r tbl-participants}
#| label: tbl-participants
#| tbl-cap: "Demographic and linguistic profile of testing sessions."
participants_bcn |> 
	unnest_wider(doe) |> 
	select(-matches("others")) |> 
	arrange(lp, test_language) |> 
	distinct(session_id, child_id, .keep_all = TRUE) |> 
	summarise(n = n(),
			  n_participants = length(unique(child_id)),
			  across(age, lst(mean, sd, min, max)),
			  across(matches("doe_"), lst(mean, sd)),
			  # age_vctr = list(age),
			  .by = c(lp, test_language)) |> 
	gt(rowname_col = "test_language",
	   groupname_col = "lp") |> 
	summary_rows(groups = c("Monolingual", "Bilingual"),
				 columns = matches("age"),
				 fns = list(Total ~ mean(.)),
				 fmt = ~ fmt_number(., decimals = 1, use_seps = FALSE)) |> 
	summary_rows(groups = c("Monolingual", "Bilingual"),
				 columns = c(n, n_participants),
				 fns = list(Total ~ sum(.)),
				 fmt = ~ fmt_integer(.)) |> 
	summary_rows(groups = c("Monolingual", "Bilingual"),
				 columns = matches("doe"),
				 fns = list(Total ~ mean(.)),
				 fmt = ~ fmt_number(., decimals = 1, scale_by = 100)) |> 
	grand_summary_rows(columns = matches("age"),
					   fns = list(Total ~ mean(.)),
					   fmt = ~ fmt_number(., decimals = 1)) |> 
	grand_summary_rows(columns = matches("doe"),
					   fns = list(Total ~ mean(.)),
					   fmt = ~ fmt_number(., decimals = 1, scale_by = 100)) |> 
	grand_summary_rows(columns = c(n, n_participants),
					   fns = list(Total ~ sum(.)),
					   fmt = ~ fmt_integer(.)) |> 
	# gt_plt_dist(age_vctr, type = "histogram") |>
	fmt_number(is.numeric, decimals = 1) |> 
	fmt_integer(c(n, n_participants)) |> 
	fmt_number(matches("doe"), decimals = 1, scale_by = 100) |> 
	cols_merge(c(age_min, age_max), 
			   pattern = "{1}-{2}") |> 
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(doe_spanish_mean, doe_spanish_sd), 
			   pattern = "{1} ({2})") |> 
	tab_spanner("Age (months)", matches("age")) |> 
	tab_spanner("Catalan", matches("catalan")) |> 
	tab_spanner("English", matches("english")) |> 
	tab_spanner("Spanish", matches("spanish")) |> 
	tab_spanner("Degree of Exposure (%)", matches("doe")) |> 
	cols_label(n = "Sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

We tested the differences between the conditions of interest in two ways. First, we examined the posterior distribution of the regression coefficients of the linear predictors in model $\mathcal{M}_0$ (see @eq-model). We assessed the practical relevance of the coefficients following @kruschke2018bayesian. We specified a region of practical equivalence (ROPE) from `r rope_interval[1]` to +`r rope_interval[2]`, in the logit scale. This region indicates the range of values that we considered equivalent to zero. We then summarised the posterior distribution of each regression coefficient with the 95% highest density interval (HDI). This interval contains the true value of this coefficient with 95% probability, given the data. Finally, we calculated the proportion of posterior samples in the 95% HDI that fell into the ROPE, noted as $p(\text{ROPE})$, which indicates the probability that the true value of the regression coefficient falls into the ROPE (and therefore should be considered equivalent to zero). For example, $p(\text{ROPE})=.80$ indicates that, given our data, there is a 80% probability that the true value of the coefficient falls within the ROPE, and can therefore be considered equivalent to zero. Second, we examined the differences between the priming conditions across the time course of the trial. We generated marginal posterior predictions from the model for the three conditions, for monolinguals and monolinguals. These predictions included the non-linear relationship between target looking and the time course of the trial. We then computed the difference between the marginal posterior predictions for the *Unrelated* and *Related/Non-cognate* conditions, and between those for the *Related/Non-cognate* and *Related/Cognate* conditions, at each time bin. Finally, we compared the 95% HDI of the differences against the \[-0.1, +0.1\] ROPE.

@@
